---
title: "homework_3"
format: html
editor: visual
---

## Problem 1.

```{r}
## reading in packages

# general use
library(tidyverse)
library(readxl)
library(here)
library(janitor)

# visualizing pairs
library(GGally)

# model selection
library(MuMIn)

# model predictions
library(ggeffects)

# model tables
library(gtsummary)
library(flextable)
library(modelsummary)

drought_exp <- read_xlsx(path = here("data", 
                                     "Valliere_etal_EcoApps_Data.xlsx"),
                         sheet = "First Harvest")

# quick look at data 
str(drought_exp)
class(drought_exp)
```

```{r}
# cleaning
drought_exp_clean <- drought_exp %>% 
  clean_names() %>% # nicer column names
  mutate(species_name = case_when( # adding column with species scientific names
    species == "ENCCAL" ~ "Encelia californica", # bush sunflower
    species == "ESCCAL" ~ "Eschscholzia californica", # California poppy
    species == "PENCEN" ~ "Penstemon centranthifolius", # Scarlet bugler
    species == "GRICAM" ~ "Grindelia camporum", # great valley gumweed
    species == "SALLEU" ~ "Salvia leucophylla", # Purple sage
    species == "STIPUL" ~ "Nasella pulchra", # Purple needlegrass
    species == "LOTSCO" ~ "Acmispon glaber" # deerweed
  )) %>% 
  relocate(species_name, .after = species) %>% # moving species_name column after species
  mutate(water_treatment = case_when( # adding column with full treatment names
    water == "WW" ~ "Well watered",
    water == "DS" ~ "Drought stressed"
  )) %>% 
  relocate(water_treatment, .after = water) # moving water_treatment column after water
```

```{r}
#visualizing and making correlation for background info
ggpairs(drought_exp_clean, # data frame
        columns = c("leaf_dry_weight_g", # columns to visualize
                    "sla", 
                    "shoot_g", 
                    "root_g", 
                    "total_g"), 
        upper = list(method = "pearson")) + # calculating Pearson correlation coefficient
  theme_bw() + # cleaner theme
  theme(panel.grid = element_blank()) # getting rid of gridlines

```

### how does total biomass differ between species?

```{r}
ggplot(data = drought_exp_clean, # data frame
       aes(x = reorder(species_name, # reordering x-axis
                       -total_g, # in reverse order of mean total mass
                       fun = mean), # calculating mean to reorder
           y = total_g)) + # y-axis
  geom_jitter(width = 0.1, # narrow jitter
              height = 0) # not jittering points up and down
```

**Plants with larger biomass seem to have larger ranges. Grindelia has highest biomass and Acrispon has lowest biomass.**

### how does total biomass differ between water treatments?

```{r}
ggplot(data = drought_exp_clean, # data frame
       aes(x = water_treatment, # x-axis
           y = total_g)) + # y-axis
  geom_jitter(width = 0.1, # narrow jitter
              height = 0) # not jittering points up and down
```

**The total biomass is lower in the drought stressed water treatment compared to the well watered treatment. Well watered plants tend to have larger biomass. This makes sense based on how plants work.**

### how does specific leaf area (SLA) influence total biomass?

```{r}
ggplot(data = drought_exp_clean, # data frame
       aes(x = sla, # x-axis
           y = total_g)) + # y-axis
  geom_point() # scatterplot
```

**Specific leaf area doesn't really effect total biomass when looking visually but it is important to check correlation to compare visual and statistical.**

## 0. Null model

```{r}
model0 <- lm(total_g ~ 1, # formula
             data = drought_exp_clean) # data frame
```

## 1. total biomass as a function of SLA, water treatment, and species

```{r}
# saturated model
model1 <- lm(total_g ~ sla + water_treatment + species_name,
             data = drought_exp_clean)

par(mfrow = c(2, 2))
plot(model1)
```

## 2. total biomass as a function of SLA and water treatment

Diagnostics for saturated model look good.

```{r}
model2 <- lm(total_g ~ sla + water_treatment,
             data = drought_exp_clean)

plot(model2)
```

## 3. total biomass as a function of SLA and species

```{r}
model3 <- lm(total_g ~ sla + species_name,
             data = drought_exp_clean)

plot(model3)
```

## 4. my own model: total biomass as a function of water treatment and species

```{r}
model4 <- lm(total_g ~ water_treatment + species_name,
             data = drought_exp_clean)

plot(model4)
```

```{r}
summary(model0)
```

```{r}
summary(model1)
```

```{r}
summary(model2)
```

```{r}
summary(model3)
```

```{r}
summary(model4)
```

## 

## A) Table of all models

```{r}
# comparing models in table
modelsummary::modelsummary( # this function takes a list of models
  list( 
    "null" = model0, # "model name" = model object
    "model 1" = model1,
    "model 2" = model2,
    "model 3" = model3,
    "model 4" = model4))
```

```{r}
model.sel(model0,
          model1, 
          model2, 
          model3,
          model4)
```

## Caption:

Table 1: Models with Predictors

This table represents the models, in each row, with their respective predictors, in each column. The right side of the table includes the comparative statistics such as the AIC, delta, and weight values.

## B) Statistical Methods

To examine the influence of the species, water treatment, and specific leaf areas on the total biomass, I created five different models comparing these variables to determine which one best describes total biomass. To determine the model that best described total biomass, I compared all of the models in a table and found the one that conformed to assumptions of linear models and had an AIC value of 0 which was model 4. To evaluate linear model assumptions, I made visualizations of each model and analyzed the homoscedasticity, normality, and outliers.

## C) Visualization

```{r}

ggpredict(model1, 
                         terms = c("sla",
                                   "water_treatment", 
                                   "species_name"))
ggpredict(model2, 
                         terms = c("sla", 
                                   "water_treatment"))
ggpredict(model3, 
                         terms = c("sla", 
                                   "species_name"))
ggpredict(model4, 
                         terms = c("water_treatment", 
                                   "species_name"))
```

```{r}
# creating new data frame of model predictions for plotting
model_preds_for_plotting <- model_preds %>% 
  rename(sla = x, # renaming columns to make this easier to use
         water_treatment = group,
         species_name = facet)

# use View(model_preds_for_plotting) 
# to compare this to the original model_preds data frame

ggplot() +
  # underlying data
  geom_point(data = drought_exp_clean,
             aes(x = sla,
                 y = total_g,
                 color = water_treatment)) +
  # model prediction 95% CI ribbon
  geom_ribbon(data = model_preds_for_plotting,
              aes(x = sla, 
                  y = predicted,
                  ymin = conf.low,
                  ymax = conf.high,
                  fill = water_treatment),
              alpha = 0.2) +
  # model prediction lines
  geom_line(data = model_preds_for_plotting,
            aes(x = sla, 
                y = predicted,
                color = water_treatment)) +
  # cleaner theme
  theme_classic() +
  # creating different panels for species
  facet_wrap(~species_name) 
```

## Caption:

Figure 1: Relationship Between Specific Leaf Area (SLA) and Total Growth Across Different Water Treatments and Species

This figure represents the relationship between SLA and total growth under different water treatments with separate panels for each species. The underlying data points are represented by blue, green, and red dots for the different water treatments. The lines represent model predictions, and the shaded areas around these lines show the 95% confidence intervals for the predictions. The visualization highlights how SLA and water treatment affect total growth for each species, with the confidence intervals indicating the uncertainty around the predictions. Data source: Valliere, Justin; Zhang, Jacqueline; Sharifi, M.; Rundel, Philip (2019). Data from: Can we condition native plants to increase drought tolerance and improve restoration success? \[Dataset\]. Dryad. <https://doi.org/10.5061/dryad.v0861f7>

## E) Results

In this analysis, I found that the predictors that best described the total biomass were the water treatment and species. As shown in the model 4 summary, the residual standard error is 0.072 on 62 degrees of freedom, an R-squared value of 0.72, F-statistic of 27.08, a p-value of \<0.001, an AIC of 0, and visually it conformed to assumptions of linear models. On average, the well watered treatments had a larger total mass across all species. When examining the differences between species, I found notable variations in total biomass between Acmispon glaber exhibiting the lowest average mass and Grindelia camporum showing the highest. These findings underscore the importance of considering both environmental factors, such as water treatment, and biological factors, such as species differences, when evaluating total mass in this analysis.

## Problem 2.

## A) 

An affective visualization for my personal data could take the form of a dynamic, interactive family tree. Each branch or node of the tree represents a family member, and the size, color, or movement of the branches reflects the frequency and duration of phone calls with each individual. This affective visualization will aim to convey the depth and complexity of familial relationships through the lens of phone conversations, fostering emotional connections and prompting the viewers to reflect on the significance of their interactions with family members over time. I think this could also work as a series of drawings, the first being the entire tree, then a page that focuses on each branch and include all the details from my data for each individual.

## B)

ADD PHOTO OF SKETCH HERE

## C)

ADD PHOTO OF DRAFT HERE

## D)

In my piece, I explore the dynamic of my cellular connections with my family members through a dynamic family tree. My first draft is a series of digital drawings where I first show each family member as their own branch on a large tree. Each branch is unique to the specific data for that individual including total duration I have spent on the phone with them, the general mood I was in speaking with them, and the frequency of which I spoke to them. I decided to do a digital art piece because I thought that would be the most effective way to visually represent my data in a way that others can connect with. The following drawings are of each branch close up to offer more details for each individual. I was influenced by the beautiful trees in my backyard at home, techniques of digital art that I found when exploring the internet, and a color scheme that I was exposed to while outside drawing. I created my first draft on my IPad using an app called Goodnotes and experimented with various pen styles and added photos of my family members from my camera roll.

## Problem 3.
